---
title: "Crop data analyses"
author: "Roy Sanderson"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("multi_plot.R")
```

## Background
Analyses for the "CHN 2020 STC data including N rates.xlsx" spreadsheet separated by year (2018 or 2019) and use farm as a random effect. Multiple comparison tests also included to show differences between management treatments.

For the data in "CHN 2020 STC data including N rates.xlsx" the design is vaguely like a randomised-block design, in that each column of 3 plots has all three management types randomly assigned. However, the nitrogen treatments are applied in a line. Thus, if the analysis is done separately, for each N treatment, it could be analysed as a randomised block design. However, I'm assuming that both N and management regime (bio, ipm, conv) are of interest. As such, this is not a true randomised block design, and therefore this has been analysed as a fully-randomised setup with binomial GLM (response are percentage disease). The growth stage could be used as either a fixed- or random-effect (for the latter, a categorical class). I've chosen to use a fixed effect, as there were in some plots trends of disease going up or down over time (all interactions initially fitted). However, in practice this proved not significant, although would need to explore the data further. As the last growth stage (GS89, 30th July 2020) merely records height, this has been ignored.

```{r crop data 2018 and 2019, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)

crop_data_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                              sheet = "Crop data 2018",
                              skip = 4,
                              na = "-")
colnames(crop_data_2018) <- c("farm", "treatment", "variety", "yield",
                              "sp_wght", "moisture", "N", "protein", "GLA")
crop_data_2018$farm[crop_data_2018$farm == 1] <- "stc"
crop_data_2018$farm[crop_data_2018$farm == 2] <- "nafferton"
crop_data_2018$farm[crop_data_2018$farm == 3] <- "cockle"
crop_data_2018$farm <- as.factor(crop_data_2018$farm)
crop_data_2018$treatment[crop_data_2018$treatment == 1] <- "bio"
crop_data_2018$treatment[crop_data_2018$treatment == 2] <- "ipm"
crop_data_2018$treatment[crop_data_2018$treatment == 3] <- "conv"
crop_data_2018$treatment <- as.factor(crop_data_2018$treatment)
crop_data_2018$variety[crop_data_2018$variety == 1] <- "leeds"
crop_data_2018$variety[crop_data_2018$variety == 2] <- "skyfall"
crop_data_2018$variety <- as.factor(crop_data_2018$variety)

crop_data_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                              sheet = "Crop data 2019",
                              skip = 4,
                              na = "-")
colnames(crop_data_2019) <- c("farm", "treatment", "variety", "yield",
                              "sp_wght", "moisture", "N", "protein", "GLA")
crop_data_2019$farm[crop_data_2019$farm == 1] <- "stc"
crop_data_2019$farm[crop_data_2019$farm == 2] <- "cockle"
crop_data_2019$farm <- as.factor(crop_data_2019$farm)
crop_data_2019$treatment[crop_data_2019$treatment == 1] <- "bio"
crop_data_2019$treatment[crop_data_2019$treatment == 2] <- "ipm"
crop_data_2019$treatment[crop_data_2019$treatment == 3] <- "conv"
crop_data_2019$treatment <- as.factor(crop_data_2019$treatment)
crop_data_2019$variety[crop_data_2019$variety == 1] <- "sundance"
crop_data_2019$variety <- as.factor(crop_data_2019$variety)

```

```{r disease data 2018-2019, echo=FALSE, message=FALSE}
# Trickier as multiple spreadsheets and duplicate column names
# STC first
snake_case <- function(nms) tolower(gsub("[.]", "_", nms))
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2018",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2018",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "C5:L5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")


disease_stc_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease STC 2018",
                                skip = 4,
                                na = "-")
colnames(disease_stc_2018) <- c("treatment", "variety", gs_di)
disease_stc_2018$treatment[disease_stc_2018$treatment == 1] <- "bio"
disease_stc_2018$treatment[disease_stc_2018$treatment == 2] <- "ipm"
disease_stc_2018$treatment[disease_stc_2018$treatment == 3] <- "conv"
disease_stc_2018$treatment <- as.factor(disease_stc_2018$treatment)
disease_stc_2018$variety[disease_stc_2018$variety == 1] <- "leeds"
disease_stc_2018$variety[disease_stc_2018$variety == 2] <- "skyfall"
disease_stc_2018$variety <- as.factor(disease_stc_2018$variety)
disease_stc_2018$farm <- "stc"

# Format not quite the same for Nafferton and Cockle Park
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP and Naff 2018",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP and Naff 2018",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "D5:G5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_cpnf_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease CP and Naff 2018",
                                skip = 4,
                                na = "-")
colnames(disease_cpnf_2018) <- c("farm", "treatment", "variety", gs_di)
disease_cpnf_2018$farm[disease_cpnf_2018$farm == 1] <- "nafferton"
disease_cpnf_2018$farm[disease_cpnf_2018$farm == 2] <- "cockle"
disease_cpnf_2018$farm <- as.factor(disease_cpnf_2018$farm)
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 1] <- "bio"
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 2] <- "ipm"
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 3] <- "conv"
disease_cpnf_2018$treatment <- as.factor(disease_cpnf_2018$treatment)
disease_cpnf_2018$variety[disease_cpnf_2018$variety == 1] <- "leeds"
disease_cpnf_2018$variety[disease_cpnf_2018$variety == 2] <- "skyfall"
disease_cpnf_2018$variety <- as.factor(disease_cpnf_2018$variety)


# Format again different for 2019 at only 2 sites
# STC 2019 disease
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2019 (standard N)",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2019 (standard N)",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "B5:C5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_stc_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease STC 2019 (standard N)",
                                skip = 4,
                                na = "-")
colnames(disease_stc_2019) <- c("treatment", gs_di)
disease_stc_2019$farm <- "stc"
disease_stc_2019$variety <- "sundance"
disease_stc_2019$treatment[disease_stc_2019$treatment == 1] <- "bio"
disease_stc_2019$treatment[disease_stc_2019$treatment == 2] <- "ipm"
disease_stc_2019$treatment[disease_stc_2019$treatment == 3] <- "conv"
disease_stc_2019$treatment <- as.factor(disease_stc_2019$treatment)


# Cockle Park 2019 disease
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP 2019",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP 2019",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "B5:C5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_cp_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease CP 2019",
                                skip = 4,
                                na = "-")
colnames(disease_cp_2019) <- c("treatment", gs_di)
disease_cp_2019$farm <- "cp"
disease_cp_2019$variety <- "sundance" # Not stated, but taken from crop data
disease_cp_2019$treatment[disease_cp_2019$treatment == 1] <- "bio"
disease_cp_2019$treatment[disease_cp_2019$treatment == 2] <- "ipm"
disease_cp_2019$treatment[disease_cp_2019$treatment == 3] <- "conv"
disease_cp_2019$treatment <- as.factor(disease_cp_2019$treatment)


```


```{r disease tidy 2018, echo=FALSE, message=FALSE}
disease_cpnf_2018_lng <- pivot_longer(disease_cpnf_2018,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")
disease_stc_2018_lng <- pivot_longer(disease_stc_2018,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")

# Deconstruct growth stages and diseases
disease_cpnf_2018_lng <- disease_cpnf_2018_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2018) %>% 
  select(-gsdis)

disease_stc_2018_lng <- disease_stc_2018_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2018) %>% 
  select(-gsdis)
       
disease_2018 <- rbind(disease_cpnf_2018_lng, disease_stc_2018_lng)
```



```{r disease tidy 2019, echo=FALSE, message=FALSE}
disease_cp_2019_lng <- pivot_longer(disease_cp_2019,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")
disease_stc_2019_lng <- pivot_longer(disease_stc_2019,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")

# Deconstruct growth stages and diseases
disease_cp_2019_lng <- disease_cp_2019_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 3, 4)), # Not the same as before
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2019) %>% 
  select(-gsdis)

disease_stc_2019_lng <- disease_stc_2019_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2019) %>% 
  select(-gsdis)
       
disease_2019 <- rbind(disease_cp_2019_lng, disease_stc_2019_lng)
```



```{r merge disease years, echo=FALSE, message=FALSE}
disease <- rbind(disease_2018, disease_2019)
```



## Basic crop graphs
Check on patterns in crops

```{r crop graphs basics, echo=FALSE, message=FALSE}
theme_set(theme_classic())
p1 <- ggplot(crop_data_2018, aes(x = farm, y=yield, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Yield 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=yield, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Yield 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=sp_wght, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Specific weight 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=sp_wght, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Specific weight 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=moisture, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain moisture content 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=moisture, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain moisture content 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=N, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain nitrogen content 2018")
p2 <- filter(crop_data_2019, farm == "stc") %>% 
  ggplot(aes(x = farm, y=N, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain nitrogen content 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=protein, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain protein content 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=protein, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain protein content 2019")
multiplot(p1, p2, cols=2)


```

## Linear mixed models
Treat farm as a random variable, as presumably it is treatment effects that are of interest. R by default does a sequential analysis. Hence, in the output of the results below the `(Intercept)` estimate is actually the mean value for the biological control `bio`. It treats this as the baseline simply because `b` is first in the alphabet. The estimate values labelled as `treatmentconv` and `treatmentipm` actually show the differences from the baseline biological control value. Hence the p-value for `(Intercept)` is to test whether the biological control differs from zero (which it always does, hence highly significant). The other two p-values show whether the differences for conventional and IPM from the biological are significant. This is a little confusing, hence I've added the multiple comparison tests.

Strictly-speaking an Analysis of Deviance table is better for a mixed-effect model than an ANOVA table, but in practical terms for a simple example like this, with only one fixed-effect, an ANOVA table is fine. The `(Intercept)` estimate is merely the overall grand mean, so can be ignored, it is the line marked as `treatment` that matters.

### Yield
#### Yield 2018
```{r mixed models, echo=FALSE, warning=FALSE, message=FALSE}
library(nlme)
library(multcomp)
library(effects)

yield_2018_lme <- lme(yield ~ treatment, random = ~ 1 | farm, data=crop_data_2018)
summary(yield_2018_lme)
```

So here, the overall yield for biological control is `r as.numeric(summary(yield_2018_lme)$coefficients$fixed[1])` and the difference for the conventional is `r as.numeric(summary(yield_2018_lme)$coefficients$fixed[2])`. Hence the overall yield for the conventional is `r as.numeric(summary(yield_2018_lme)$coefficients$fixed[1]) + as.numeric(summary(yield_2018_lme)$coefficients$fixed[2])` and similarly for IPM, which differs from the biological by `r as.numeric(summary(yield_2018_lme)$coefficients$fixed[3])` thus the IPM yield is `r as.numeric(summary(yield_2018_lme)$coefficients$fixed[1]) + as.numeric(summary(yield_2018_lme)$coefficients$fixed[3])`

```{r mixed models2, echo=FALSE, warning=FALSE, message=FALSE}
anova(yield_2018_lme)
summary(glht(yield_2018_lme, linfct = mcp(treatment = "Tukey")))
#cints <- (confint(glht(yield_2018_lme, linfct = mcp(treatment = "Tukey"))))[[10]]
#data.frame(cints) # Neater printing
#plot(allEffects(yield_2018_lme), lines=list(col="black"))
```

#### Yield 2019
```{r echo=FALSE, warning=FALSE, message=FALSE}
yield_2019_lme <- lme(yield ~ treatment, random = ~ 1 | farm, data=crop_data_2019)
summary(yield_2019_lme)
anova(yield_2019_lme)
summary(glht(yield_2019_lme, linfct = mcp(treatment = "Tukey")))
#cints <- (confint(glht(yield_2019_lme, linfct = mcp(treatment = "Tukey"))))[10]
#data.frame(cints)
```

#### Specific weight 2018
```{r echo=FALSE, warning=FALSE, message=FALSE}

lme_summary <- function(fix_form, data) {
  res_lme <- lme(fix_form, random = ~ 1 | farm, data=data)
  print(summary(res_lme))
  print(anova(res_lme))
  print(summary(glht(res_lme, linfct = mcp(treatment = "Tukey"))))
#  cints <- (confint(glht(res_lme, linfct = mcp(treatment = "Tukey"))))[10]
#  data.frame(cints)
}

lme_summary(formula(sp_wght ~ treatment), data=crop_data_2018)
```

#### Specific weight 2019
```{r echo=FALSE}
lme_summary(formula(sp_wght ~ treatment), data=crop_data_2019)
```

#### Moisture 2018
```{r echo=FALSE}
lme_summary(formula(moisture ~ treatment), data=crop_data_2018)
```

#### Moisture 2019
```{r echo=FALSE}
lme_summary(formula(moisture ~ treatment), data=crop_data_2019)
```

#### Nitrogen 2018
```{r echo=FALSE}
lme_summary(formula(N ~ treatment), data=crop_data_2018)
```

#### Nitrogen 2019
Only available for STC in 2019 therefore linear model rather than mixed-effects model.
```{r echo=FALSE}
#lme_summary(formula(N ~ treatment), data=crop_data_2019)
n_2019_lm <- lm(N ~ treatment, data=crop_data_2019[1:15,]) # only stc
summary(n_2019_lm)
summary(glht(n_2019_lm, linfct = mcp(treatment = "Tukey")))
#cints <- (confint(glht(n_2019_lm, linfct = mcp(treatment = "Tukey"))))[10]
#data.frame(cints)
```

#### Protein 2018
```{r echo=FALSE}
lme_summary(formula(protein ~ treatment), data=crop_data_2018)
```

#### Protein 2019
```{r echo=FALSE}
lme_summary(formula(protein ~ treatment), data=crop_data_2019)
```

## STC 2020 data analysis
This is for the data in the "CHN 2020 STC data including N rates.xlsx" spreadsheet.

```{r read in STC data for 2020 block expt, echo=FALSE, message=FALSE}
yield_quality <- read_excel("data/CHN 2020 STC data including N rates.xlsx",
                              sheet = "YIELD & QUALITY",
                              skip = 6,
                              range = "A7:Q43",
                              na = "-")
colnames(yield_quality) <- c("plot", "treatment", "N", "yield", "lodging", "brackling",
                             "moisture", "dm", "sp_wght", "tgw", "n_content",
                             "protein_content", "grain_2.8", "grain_2.5_2.8",
                             "grain_2.2_2.5", "grain_2.0_2.2", "grain_2.0")
# Need to allocate block numbers to plots; These are not 'true' blocks as they
# are not independent of N applications, so probably best ignored for analysis
yield_quality$block <- NA
yield_quality$block[yield_quality$plot %in% 1:3] <- "1"
yield_quality$block[yield_quality$plot %in% 4:6] <- "2"
yield_quality$block[yield_quality$plot %in% 7:9] <- "3"
yield_quality$block[yield_quality$plot %in% 10:12] <- "4"
yield_quality$block[yield_quality$plot %in% 13:15] <- "1"
yield_quality$block[yield_quality$plot %in% 16:18] <- "2"
yield_quality$block[yield_quality$plot %in% 19:21] <- "3"
yield_quality$block[yield_quality$plot %in% 22:24] <- "4"
yield_quality$block[yield_quality$plot %in% 25:27] <- "1"
yield_quality$block[yield_quality$plot %in% 28:30] <- "2"
yield_quality$block[yield_quality$plot %in% 31:33] <- "3"
yield_quality$block[yield_quality$plot %in% 34:36] <- "4"
yield_quality$block <- as.factor(yield_quality$block)

yield_quality$treatment[yield_quality$treatment == 1] <- "bio"
yield_quality$treatment[yield_quality$treatment == 2] <- "ipm"
yield_quality$treatment[yield_quality$treatment == 3] <- "conv"
yield_quality$treatment <- as.factor(yield_quality$treatment)
yield_quality$N <- as.factor(yield_quality$N) # Excel unclear on exact amounts

# Disease data for 2020
# Growth stages
gs_names <- read_excel("data/CHN 2020 STC data including N rates.xlsx",
                              sheet = "DISEASE ASSESSMENTS",
                              skip = 4,
                              range = "D5:W5",
                              na = "-")
gs_names <- colnames(gs_names)
curr_name <- gs_names[1]
for (i in 1:length(gs_names)){
  if(str_sub(gs_names[i], 1, 3) == "..."){
    gs_names[i] <- curr_name
  } else {
    curr_name <- gs_names[i]
  }
}
str_sub(gs_names, 3, 3) <- ""

# Dates
date_names <- read_excel("data/CHN 2020 STC data including N rates.xlsx",
                              sheet = "DISEASE ASSESSMENTS",
                              skip = 5,
                              range = "D6:W6",
                              na = "-")
date_names <- colnames(date_names)
date_name <- date_names[1]
for (i in 1:length(date_names)){
  if(str_sub(date_names[i], 1, 3) == "..."){
    date_names[i] <- curr_name
  } else {
    curr_name <- date_names[i]
  }
}
str_sub(date_names, 3, 3) <- ""
date_names <- as.Date(as.numeric(date_names), origin="2008-06-01") # origin trial and error

# Actual disease data
disease_raw <- read_excel("data/CHN 2020 STC data including N rates.xlsx",
                          sheet = "DISEASE ASSESSMENTS",
                          skip = 6,
                          na = "-",
                          .name_repair = "universal")
di <- snake_case(colnames(disease_raw) <- str_to_lower(colnames(disease_raw)))
di <- str_replace_all(di, "_____", "")
di <- str_replace_all(di, "____", "")
di <- str_replace_all(di, "__", "")
di <- str_replace_all(di, "_", "")

da_gs_di <- str_c(date_names, gs_names, di[-(1:3)])

colnames(disease_raw) <- c(di[1:3], da_gs_di)

# Convert to long format
disease_lng <- pivot_longer(disease_raw,
                            cols = 4:23,
                            names_to = "health",
                            values_to = "pcnt")
disease_lng$date <- as.Date(str_sub(disease_lng$health, 1, 10))
disease_lng$gs   <- str_sub(disease_lng$health, 11, 14)
disease_lng$health <- str_extract(disease_lng$health, "septoria|mildew|yellowrust|flaggla|greenleafarea")
colnames(disease_lng)[3] <- "N"  # be consistent with earlier
disease_lng$N <- as.factor(disease_lng$N)
disease_lng$treatment[disease_lng$treatment == 1] <- "bio"
disease_lng$treatment[disease_lng$treatment == 2] <- "ipm"
disease_lng$treatment[disease_lng$treatment == 3] <- "conv"

# Not interested in GS89 as only contains height
disease_lng <- filter(disease_lng, gs != "GS89")
```

### Broad patterns in crop yield and quality
General patterns for STC 2020 crop data. Lodging figures not shown as nearly all values are zero. Nitrogen has been coded as 1, 2 or 3 as actual applications in spreadsheet do not agree between the "TRIAL LAYOUT" vs the "YIELD & QUALITY" or "DISEASE ASSESSMENT" tabs.

```{r, echo = FALSE}
p1 <- ggplot(yield_quality, aes(x = N, y=yield, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain yield (t/ha)")
p2 <- ggplot(yield_quality, aes(x = N, y=lodging, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Root lodging (%)")
p3 <- ggplot(yield_quality, aes(x = N, y=brackling, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Brackling (%)")
p4 <- ggplot(yield_quality, aes(x = N, y=moisture, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Moisture (%)")
p5 <- ggplot(yield_quality, aes(x = N, y=dm, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Dry matter (%)")

multiplot(p1, p3, p4, p5, cols=2)
```

General patterns for some additional information on grain chemistry collected by NIAB.

```{r, echo = FALSE}
p1 <- ggplot(yield_quality, aes(x = N, y=sp_wght, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Specific weight (kg/hl)")
p2 <- ggplot(yield_quality, aes(x = N, y=tgw, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Thousand grain weight (g)")
p3 <- ggplot(yield_quality, aes(x = N, y=n_content, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Nitrogen content (g/100g)")
p4 <- ggplot(yield_quality, aes(x = N, y=protein_content, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Protein content (g/100g)")

multiplot(p1, p2, p3, p4, cols=2)
```

Physical characteristics of harvested grain measured by NIAB

```{r, echo = FALSE}
p1 <- ggplot(yield_quality, aes(x = N, y=grain_2.8, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain >2.8 mm")
p2 <- ggplot(yield_quality, aes(x = N, y=grain_2.5_2.8, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain 2.5-2.8 mm")
p3 <- ggplot(yield_quality, aes(x = N, y=grain_2.2_2.5, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain 2.2-2.5 mm")
p4 <- ggplot(yield_quality, aes(x = N, y=grain_2.0_2.2, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain 2.0-2.2 mm")
p5 <- ggplot(yield_quality, aes(x = N, y=grain_2.0, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain <2.0 mm")

multiplot(p1, p3, p5, p2, p4, cols=2)
```

### Broad patterns in disease at STC 2020
These data could be plotted for each growth stage, although that would be quite over-whelming in terms of numbers of graphs. These are for GS71, i.e. 10th July 2020.

```{r, echo=FALSE, message=FALSE}
p1 <- filter(disease_lng, health == "septoria" & gs == "GS71") %>% 
  ggplot(aes(x = N, y=pcnt, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Septoria (%)")
p2 <- filter(disease_lng, health == "yellowrust" & gs == "GS71") %>% 
  ggplot(aes(x = N, y=pcnt, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Yellow rust (%)")
p3 <- filter(disease_lng, health == "mildew" & gs == "GS71") %>% 
  ggplot(aes(x = N, y=pcnt, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Mildew (%)")
p4 <- filter(disease_lng, health == "greenleafarea" & gs == "GS71") %>% 
  ggplot(aes(x = N, y=pcnt, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Green leaf area (%)")

multiplot(p1, p2, p3, p4, cols=2)
```

Changes over time for Septoria

```{r, message=FALSE, echo=FALSE}
tmp <- filter(disease_lng, health == "septoria") %>%
  group_by(treatment, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p1 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=treatment)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Septoria (%)")

tmp <- filter(disease_lng, health == "septoria") %>%
  group_by(N, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p2 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=N)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Septoria (%)")
multiplot(p1, p2, cols=2)
```

Changes over time for yellow rust

```{r, message=FALSE, echo=FALSE}
tmp <- filter(disease_lng, health == "yellowrust") %>%
  group_by(treatment, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p1 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=treatment)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Yellow rust (%)")

tmp <- filter(disease_lng, health == "yellowrust") %>%
  group_by(N, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p2 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=N)) +
  geom_point() +
  geom_line()  + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Yellow rust (%)")
multiplot(p1, p2, cols=2)
```

Changes over time for mildew

```{r, message=FALSE, echo=FALSE}
tmp <- filter(disease_lng, health == "mildew") %>%
  group_by(treatment, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p1 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=treatment)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Mildew (%)")

tmp <- filter(disease_lng, health == "yellowrust") %>%
  group_by(N, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p2 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=N)) +
  geom_point() +
  geom_line()  + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Mildew (%)")
multiplot(p1, p2, cols=2)
```

Changes over time for green leaf area

```{r, message=FALSE, echo=FALSE}
tmp <- filter(disease_lng, health == "greenleafarea") %>%
  group_by(treatment, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p1 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=treatment)) +
  geom_point() +
  geom_line() + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Green leaf area (%)")

tmp <- filter(disease_lng, health == "greenleafarea") %>%
  group_by(N, date) %>% 
  summarise(
    meanpcnt = mean(pcnt, na.rm = TRUE),
    sepcnt = sd(pcnt, na.rm = TRUE)/sqrt(n())
  )
  
p2 <- ggplot(tmp, aes(x = date, y = meanpcnt, colour=N)) +
  geom_point() +
  geom_line()  + 
  geom_errorbar(aes(ymin = meanpcnt - sepcnt,
                    ymax = meanpcnt + sepcnt,
                    width = 0.5)) +
  ylab("Green leaf area (%)")
multiplot(p1, p2, cols=2)
```

### Generalised linear models of disease at STC
These data are not ideal for analysis due to the pseudo-blocking setup. So mixed-effects models don't really work properly here, and not ideal for a repeated-measures model. To get round the problem, a full binomial GLM model with 3 main effects + 2-level interactions + 3-level interaction is created first, then gradually simplifed.

**Septoria** No significant treatment, N or date effects. No significant higher-level interactions (not shown).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)
disease_lng <- disease_lng[!is.na(disease_lng$pcnt),]
disease_lng$plotno <- as.factor(disease_lng$plotno)
disease_lng <- mutate(disease_lng, prop = pcnt/100)
disease_lng <- mutate(disease_lng, dayno = yday(date))

septoria <- filter(disease_lng, health == "septoria")
septoria_glm1 <- glm(prop ~ treatment * N * dayno, data=septoria, family="binomial")
septoria_glm2 <- update(septoria_glm1, . ~ . - treatment:N:dayno)
#anova(septoria_glm1, septoria_glm2, test="Chi")
septoria_glm3 <- update(septoria_glm2, . ~ . - treatment:dayno)
#anova(septoria_glm2, septoria_glm3, test="Chi")
septoria_glm4 <- update(septoria_glm3, . ~ . - N:dayno)
#anova(septoria_glm3, septoria_glm4, test="Chi")
septoria_glm5 <- update(septoria_glm4, . ~ . - treatment:N)
#anova(septoria_glm4, septoria_glm5, test="Chi")
summary(septoria_glm5)
```

**Mildew** No significant treatment, N or date effects. No significant higher-level interactions (not shown).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mildew <- filter(disease_lng, health == "mildew")
mildew_glm1 <- glm(prop ~ treatment * N * dayno, data=mildew, family="binomial")
mildew_glm2 <- update(mildew_glm1, . ~ . - treatment:N:dayno)
#anova(mildew_glm1, mildew_glm2, test="Chi")
mildew_glm3 <- update(mildew_glm2, . ~ . - treatment:dayno)
#anova(mildew_glm2, mildew_glm3, test="Chi")
mildew_glm4 <- update(mildew_glm3, . ~ . - N:dayno)
#anova(mildew_glm3, mildew_glm4, test="Chi")
mildew_glm5 <- update(mildew_glm4, . ~ . - treatment:N)
#anova(mildew_glm4, mildew_glm5, test="Chi")
summary(mildew_glm5)
```


**Yellow rust** No significant treatment, N or date effects. No significant higher-level interactions (not shown).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
yellowrust <- filter(disease_lng, health == "yellowrust")
yellowrust_glm1 <- glm(prop ~ treatment * N * dayno, data=yellowrust, family="binomial")
yellowrust_glm2 <- update(yellowrust_glm1, . ~ . - treatment:N:dayno)
#anova(yellowrust_glm1, yellowrust_glm2, test="Chi")
yellowrust_glm3 <- update(yellowrust_glm2, . ~ . - treatment:dayno)
#anova(yellowrust_glm2, yellowrust_glm3, test="Chi")
yellowrust_glm4 <- update(yellowrust_glm3, . ~ . - N:dayno)
#anova(yellowrust_glm3, yellowrust_glm4, test="Chi")
yellowrust_glm5 <- update(yellowrust_glm4, . ~ . - treatment:N)
#anova(yellowrust_glm4, yellowrust_glm5, test="Chi")
summary(yellowrust_glm5)
```

### STC 2020 crop data
Various analyses for yield, specific weight, thousand grain weight, nitrogen content etc. Note that nitrogen application `N` is coded as a factor, rather than a continuous variable, as the spreadsheet is inconsistent in the actual amounts applied. When it is coded as a continuous numeric variable, there are more degrees of freedom for error (F~5,30~ rather than F~8,27~) as fewer coefficients need to be estimated. This has an effect on the p-values, i.e. some are significant if N is continuous (coded as numbers 1, 2, 3) but non-significant as factors (N with levels "1", "2", "3"). Ideally the true N values as continuous numeric variables (120, 140 etc.) would be used in the analysis rather than 1,2,3.

**Yield**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# As these are balanced designs, no need for car::Anova as results the same
yield_lm  <- lm(yield ~ treatment * N, data=yield_quality)
yield_lm2 <- lm(yield ~ treatment * as.numeric(N), data=yield_quality)
anova(yield_lm)
anova(yield_lm2)
```

**Dry matter**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
dm_lm  <- lm(dm ~ treatment * N, data=yield_quality)
dm_lm2 <- lm(dm ~ treatment * as.numeric(N), data=yield_quality)
anova(dm_lm)
anova(dm_lm2)
```

There appears to be a significant effect of N application on dry matter, with as you might expect, higher dry matter with higher N. The treatment:N interactions probably reflects the fact that at low N the conventional management gives the highest DM, whereas at high N the conventional management gives the lowest DM of the three regimes.

**Specific weight**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
sp_wght_lm  <- lm(sp_wght ~ treatment * N, data=yield_quality)
sp_wght_lm2 <- lm(sp_wght ~ treatment * as.numeric(N), data=yield_quality)
anova(sp_wght_lm)
anova(sp_wght_lm2)
```

**Thousand grain weight**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
tgw_lm  <- lm(tgw ~ treatment * N, data=yield_quality)
tgw_lm2 <- lm(tgw ~ treatment * as.numeric(N), data=yield_quality)
anova(tgw_lm)
anova(tgw_lm2)
```

**Nitrogen content**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
n_content_lm  <- lm(n_content ~ treatment * N, data=yield_quality)
n_content_lm2 <- lm(n_content ~ treatment * as.numeric(N), data=yield_quality)
anova(n_content_lm)
anova(n_content_lm2)
```

Obvious differences in nitrogen content. Post-hoc tests (with N application considered as continuous numeric 1,2,3).

```{r, warning=FALSE, message=FALSE, echo=FALSE}
summary(n_content_lm2)
summary(glht(n_content_lm2, linfct = mcp(treatment = "Tukey")))
```

As you would expect, grain nitrogen content is highest with the highest N application rate. The significant interaction term reflects the fact that at low and medium N application the IPM management has lowest grain nitrogen, whereas at high N application the IPM has highest grain nitrogen content.

**Protein content**

```{r, warning=FALSE, message=FALSE, echo=FALSE}
protein_content_lm  <- lm(protein_content ~ treatment * N, data=yield_quality)
protein_content_lm2 <- lm(protein_content ~ treatment * as.numeric(N), data=yield_quality)
anova(protein_content_lm)
anova(protein_content_lm2)
```

Again, significant differences in protein content (including both main effects and the interaction term). Multiple comparison tests on treatment (with N considered continuous numeric 1,2,3)

```{r, warning=FALSE, message=FALSE, echo=FALSE}
summary(protein_content_lm2)
summary(glht(protein_content_lm2, linfct = mcp(treatment = "Tukey")))
```

Here both main effects and the interaction terms are significant. Highest protein content with highest N application, and overall the highest protein with conventional and lowest with IPM management. However, the interaction term is significant, and at the highest N application rate the protein content is actually highest in IPM.