---
title: "Crop data analyses"
author: "Roy Sanderson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("multi_plot.R")
```

## Background
Dave needs separate analyses each year

In the Excel spreadsheet "CHN 2018 and 2019 data all sites.xlsx" under "Crop data 2018" and "Crop data 2019" columns D to H are key response variables. %N is not available for some entries. GLA is green leave area.

The other spreadsheet "CHN 2020 STC data including N rates.xlsx" has plot numbers (it actually is a randomised block design although isn't coded as such), yield and quality, and NIAB assessments.


```{r crop data 2018 and 2019, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)

crop_data_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                              sheet = "Crop data 2018",
                              skip = 4,
                              na = "-")
colnames(crop_data_2018) <- c("farm", "treatment", "variety", "yield",
                              "sp_wght", "moisture", "N", "protein", "GLA")
crop_data_2018$farm[crop_data_2018$farm == 1] <- "stc"
crop_data_2018$farm[crop_data_2018$farm == 2] <- "nafferton"
crop_data_2018$farm[crop_data_2018$farm == 3] <- "cockle"
crop_data_2018$farm <- as.factor(crop_data_2018$farm)
crop_data_2018$treatment[crop_data_2018$treatment == 1] <- "bio"
crop_data_2018$treatment[crop_data_2018$treatment == 2] <- "ipm"
crop_data_2018$treatment[crop_data_2018$treatment == 3] <- "conv"
crop_data_2018$treatment <- as.factor(crop_data_2018$treatment)
crop_data_2018$variety[crop_data_2018$variety == 1] <- "leeds"
crop_data_2018$variety[crop_data_2018$variety == 2] <- "skyfall"
crop_data_2018$variety <- as.factor(crop_data_2018$variety)

crop_data_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                              sheet = "Crop data 2019",
                              skip = 4,
                              na = "-")
colnames(crop_data_2019) <- c("farm", "treatment", "variety", "yield",
                              "sp_wght", "moisture", "N", "protein", "GLA")
crop_data_2019$farm[crop_data_2019$farm == 1] <- "stc"
crop_data_2019$farm[crop_data_2019$farm == 2] <- "cockle"
crop_data_2019$farm <- as.factor(crop_data_2019$farm)
crop_data_2019$treatment[crop_data_2019$treatment == 1] <- "bio"
crop_data_2019$treatment[crop_data_2019$treatment == 2] <- "ipm"
crop_data_2019$treatment[crop_data_2019$treatment == 3] <- "conv"
crop_data_2019$treatment <- as.factor(crop_data_2019$treatment)
crop_data_2019$variety[crop_data_2019$variety == 1] <- "sundance"
crop_data_2019$variety <- as.factor(crop_data_2019$variety)

```

```{r disease data 2018-2019, echo=FALSE, message=FALSE}
# Trickier as multiple spreadsheets and duplicate column names
# STC first
snake_case <- function(nms) tolower(gsub("[.]", "_", nms))
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2018",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2018",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "C5:L5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")


disease_stc_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease STC 2018",
                                skip = 4,
                                na = "-")
colnames(disease_stc_2018) <- c("treatment", "variety", gs_di)
disease_stc_2018$treatment[disease_stc_2018$treatment == 1] <- "bio"
disease_stc_2018$treatment[disease_stc_2018$treatment == 2] <- "ipm"
disease_stc_2018$treatment[disease_stc_2018$treatment == 3] <- "conv"
disease_stc_2018$treatment <- as.factor(disease_stc_2018$treatment)
disease_stc_2018$variety[disease_stc_2018$variety == 1] <- "leeds"
disease_stc_2018$variety[disease_stc_2018$variety == 2] <- "skyfall"
disease_stc_2018$variety <- as.factor(disease_stc_2018$variety)
disease_stc_2018$farm <- "stc"

# Format not quite the same for Nafferton and Cockle Park
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP and Naff 2018",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP and Naff 2018",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "D5:G5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_cpnf_2018 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease CP and Naff 2018",
                                skip = 4,
                                na = "-")
colnames(disease_cpnf_2018) <- c("farm", "treatment", "variety", gs_di)
disease_cpnf_2018$farm[disease_cpnf_2018$farm == 1] <- "nafferton"
disease_cpnf_2018$farm[disease_cpnf_2018$farm == 2] <- "cockle"
disease_cpnf_2018$farm <- as.factor(disease_cpnf_2018$farm)
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 1] <- "bio"
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 2] <- "ipm"
disease_cpnf_2018$treatment[disease_cpnf_2018$treatment == 3] <- "conv"
disease_cpnf_2018$treatment <- as.factor(disease_cpnf_2018$treatment)
disease_cpnf_2018$variety[disease_cpnf_2018$variety == 1] <- "leeds"
disease_cpnf_2018$variety[disease_cpnf_2018$variety == 2] <- "skyfall"
disease_cpnf_2018$variety <- as.factor(disease_cpnf_2018$variety)


# Format again different for 2019 at only 2 sites
# STC 2019 disease
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2019 (standard N)",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease STC 2019 (standard N)",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "B5:C5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_stc_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease STC 2019 (standard N)",
                                skip = 4,
                                na = "-")
colnames(disease_stc_2019) <- c("treatment", gs_di)
disease_stc_2019$farm <- "stc"
disease_stc_2019$variety <- "sundance"
disease_stc_2019$treatment[disease_stc_2019$treatment == 1] <- "bio"
disease_stc_2019$treatment[disease_stc_2019$treatment == 2] <- "ipm"
disease_stc_2019$treatment[disease_stc_2019$treatment == 3] <- "conv"
disease_stc_2019$treatment <- as.factor(disease_stc_2019$treatment)


# Cockle Park 2019 disease
gs <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP 2019",
                           skip = 3,
                           n_max = 0,
                           .name_repair = "universal"
                  ))
gs <- snake_case(gs)
gs <- str_replace_all(gs, "___", "_")
di <-  colnames(read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                           sheet = "Disease CP 2019",
                           skip = 4,
                           n_max = 0,
                           .name_repair = "universal",
                           range = "B5:C5"
                  ))
di <- snake_case(di)
di <- str_replace_all(di, "___", "_")
di <- str_replace_all(di, "__", "")
gs_di <- str_c(gs, di, sep="_")
disease_cp_2019 <- read_excel("data/CHN 2018 and 2019 data all sites.xlsx",
                                sheet = "Disease CP 2019",
                                skip = 4,
                                na = "-")
colnames(disease_cp_2019) <- c("treatment", gs_di)
disease_cp_2019$farm <- "cp"
disease_cp_2019$variety <- "sundance" # Not stated, but taken from crop data
disease_cp_2019$treatment[disease_cp_2019$treatment == 1] <- "bio"
disease_cp_2019$treatment[disease_cp_2019$treatment == 2] <- "ipm"
disease_cp_2019$treatment[disease_cp_2019$treatment == 3] <- "conv"
disease_cp_2019$treatment <- as.factor(disease_cp_2019$treatment)


```

Data are a bit of a mess, especially disease. Probably sensible to convert everything to tidy format to make easier to manage.

```{r disease tidy 2018, echo=FALSE, message=FALSE}
disease_cpnf_2018_lng <- pivot_longer(disease_cpnf_2018,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")
disease_stc_2018_lng <- pivot_longer(disease_stc_2018,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")

# Deconstruct growth stages and diseases
disease_cpnf_2018_lng <- disease_cpnf_2018_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2018) %>% 
  select(-gsdis)

disease_stc_2018_lng <- disease_stc_2018_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2018) %>% 
  select(-gsdis)
       
disease_2018 <- rbind(disease_cpnf_2018_lng, disease_stc_2018_lng)
```

Do the same for 2019

```{r disease tidy 2019, echo=FALSE, message=FALSE}
disease_cp_2019_lng <- pivot_longer(disease_cp_2019,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")
disease_stc_2019_lng <- pivot_longer(disease_stc_2019,
                                      cols = starts_with("gs"),
                                      names_to = "gsdis",
                                      values_to = "severity")

# Deconstruct growth stages and diseases
disease_cp_2019_lng <- disease_cp_2019_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 3, 4)), # Not the same as before
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2019) %>% 
  select(-gsdis)

disease_stc_2019_lng <- disease_stc_2019_lng %>%
  mutate(gs = as.numeric(str_sub(gsdis, 4, 5)),
         pathogen = str_extract(gsdis, "mildew|septoria"),
         year = 2019) %>% 
  select(-gsdis)
       
disease_2019 <- rbind(disease_cp_2019_lng, disease_stc_2019_lng)
```

May as well merge into one data.frame for both years if needed

```{r merge disease years, echo=FALSE, message=FALSE}
disease <- rbind(disease_2018, disease_2019)
```



## Basic crop graphs
Check on patterns in crops

```{r crop graphs basics, echo=FALSE, message=FALSE}
theme_set(theme_classic())
p1 <- ggplot(crop_data_2018, aes(x = farm, y=yield, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Yield 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=yield, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Yield 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=sp_wght, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Specific weight 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=sp_wght, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Specific weight 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=moisture, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain moisture content 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=moisture, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain moisture content 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=N, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain nitrogen content 2018")
p2 <- filter(crop_data_2019, farm == "stc") %>% 
  ggplot(aes(x = farm, y=N, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain nitrogen content 2019")
multiplot(p1, p2, cols=2)

p1 <- ggplot(crop_data_2018, aes(x = farm, y=protein, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain protein content 2018")
p2 <- ggplot(crop_data_2019, aes(x = farm, y=protein, colour=treatment)) +
  geom_boxplot() +
  ggtitle("Grain protein content 2019")
multiplot(p1, p2, cols=2)


```

## Linear mixed models
Treat farm as a random variable, as presumably it is treatment effects that are of interest.

```{r mixed models}
library(nlme)
library(multcomp)
library(effects)
yield_2018_lme <- lme(yield ~ treatment, random = ~ 1 | farm, data=crop_data_2018)
summary(yield_2018_lme)
summary(glht(yield_2018_lme, linfct = mcp(treatment = "Tukey")))
confint(glht(yield_2018_lme, linfct = mcp(treatment = "Tukey")))
#plot(allEffects(yield_2018_lme), lines=list(col="black"))

```

